---
title: "Efficiency of tumopp"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Efficiency of tumopp}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, setup-hidden}
#| include: false
knitr::opts_chunk$set(fig.alt = " ")
knitr::opts_chunk$set(collapse = TRUE)
knitr::opts_chunk$set(fig.height = 6)
library(ggplot2)
theme_here = theme_bw(base_size = 12) +
  theme(panel.grid.minor = element_blank(),
        axis.ticks = element_blank(),
        strip.background = element_rect(linetype = 0))
theme_set(theme_here)
```

## Run simulation with benchmark information

```{r prepare}
#| message: false
library(tumopp)

.const = list(D = 3, C = "hex", k = 10, N = 2 ** 17, benchmark = TRUE)
.alt = list(
  L = c("const", "linear"),
  P = c("random", "mindrag"),
  d = c(0, 0.2),
  m = c(0, 2)
)
args_tbl = tumopp::make_args(alt = .alt, const = .const) |> print()
```

```{r run }
#| eval: false
set.seed(24601L)
results = tumopp::tumopp(args_tbl, graph = FALSE) |> print()

unnested = results |>
  dplyr::select(!c(population, seed)) |>
  dplyr::mutate(outdir = fs::path_file(outdir)) |>
  dplyr::select(where(\(x) dplyr::n_distinct(x) > 1L)) |>
  tidyr::unnest(benchmark) |>
  print()

df_benchmark = unnested |>
  dplyr::distinct(outdir, size, .keep_all = TRUE) |>
  dplyr::mutate(
    local = factor(local, levels = .alt[["L"]]),
    path = factor(path, levels = .alt[["P"]]),
    size = size / 1000,      # to K
    memory = maxrss / 1024,  # K to M
    utime = utime / 1000,    # ms to sec
    stime = stime / 1000,    # ms to sec
    maxrss = NULL
  ) |>
  print()

# usethis::use_data(df_benchmark, internal = TRUE, overwrite = TRUE)
```

```{r use-internal}
#| include: false
df_benchmark = tumopp:::df_benchmark # nolint: undesirable_operator_linter
```

## Memory

- Memory usage increases linearly with the increasing number of cells
  because *tumopp* is an individual-based model of tumor cells.
- Memory is also used to trace back cell lineages and write them to a text file after growth simulation.
- Cell death makes cell lineages longer given the same number of extant cells.
- The effect of cell death is more pronounced
  when the population average of birth rate is decreased with local competition (`-Llinear`).

```{r memory}
ggplot(df_benchmark) +
  aes(size, memory) +
  geom_line(linewidth = 1, color = "#337ab7") +
  facet_grid(vars(local, path), vars(delta0, rho0), label = label_both) +
  labs(x = "Size (K cells)", y = "Memory (MB)")
```

## Time

- The larger a population gets, the longer it takes
    - (`-Pmindrag`) to search for nearest empty sites from dividing cells,
    - (`-Lconst`) to push out other cells,
    - (`-Llinear`) to skip event queues of dormant (surrounded) cells.
- The combination of `-Lconst -Pmindrag` is slower than the others
  because it involves both search and push from interior cells.
- The effects of cell death and migration are smaller
  than those of local competition (`-L`) and placement of daughter cells(`-P`).

```{r time}
ggplot(df_benchmark) +
  aes(size, utime + stime) +
  geom_line(linewidth = 1, color = "#337ab7") +
  facet_grid(vars(local, path), vars(delta0, rho0), label = label_both) +
  labs(x = "Size (K cells)", y = "Time (sec)")
```
